---
title: "Rejection Sampling"
author: "Amy Chen"
date: "October 16, 2018"
output: html_document
---

###Overview
Rejection sampling is a method of drawing points from a target distribution by sampling from an easier distribution and correcting the sampling probability through the rejection of some candidates.
<br><br>
Suppose we are interested in drawing points from $f(x)$. Let $g$ denote another density from which we know how to sample and for which we can easily calculate $g(x)$. Let $e$ denote an envelope such that $e(x) = \frac{g(x)}{\alpha}\geq f(x) \forall x$ for which $f(x) > 0$ for a given constant $\alpha \leq 1$.
<br><br>
Then rejection sampling can be completed in the following steps:

1. Sample $Y$ ~ $g$
2. Sample $U$ ~ $Unif(0,1)$
3. Reject $Y$ if $U > \frac{f(Y)}{e(Y)}$, otherwise keep $Y$
4. Repeat for desired sample size
<br><br>

###Demonstration
**Suppose we would like to estimate $S= E[x^2]$ where $X$ has density proportional to $q(x) = e^ \frac{-|x|^3}{3}\ $**
<br><br>

####Target Function
Let the target function be $f(x) = e^ \frac{-|x|^3}{3}\ $    
<br>

####Envelope Function
Consider using the standard normal distribution $N(0,1)$  
<center>$g = \frac{1}{\sqrt{2\pi}}e^{\frac{-x^2}{2}}$</center>  
To determine $e$, we must find $\alpha$ such that $\frac{g(x)}{\alpha} > f(x)$. Choose $\alpha = 0.3$.

```{r}
f <- function(x) {         
  exp(-abs(x^3)/3)
}
e <- function(x) {     
  dnorm(x)/0.3
}
```

```{r}
curve(e(x), from = -2, to = 2)
curve(f(x), add = T, col = "blue")
```

####Rejection Sampling
Follow the steps listed above for a sample size of 100,000. Below are a few points drawn using this method.
```{r}
n = 100000
y = rnorm(n, 0, 1)
u = runif(n)
x = y[u < f(y)/e(y)]
head(x)
```

The acceptance ratio is
```{r}
length(x)/n
```

<br>

####Calculate $E[x^2]$
Now that we have our points from the sample, square each accepted x, and take the mean to get $E[x^2]$.
```{r}
mean(x^2)
```

